---
title: "Calculating LSD Values of ASReml-R Model"
author: "Zhanglong"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(igraph)
library(asreml)
library(asremlPlus)
library(ggplot2)
library(agricolae)
library(tidyverse)
library(gridExtra)
```

# LSDlabels Function
The ```LSDlabels``` function is used for calculating the LSD values for the pairwise comparisons from the ```ASReml-R``` model.

```{r,warning=FALSE}
LSDlabels <- function(P, alpha) {
  a <- P
  diag(a) <- 1
  b <- which(a > alpha, arr.ind = TRUE)
  top <- data.frame(N1 = b[, 1], N2 = b[, 2])
  g3 <- igraph::simplify(igraph::graph.data.frame(top[order(top[[1]]), ], directed = FALSE))
  cliq <- igraph::maximal.cliques(g3)
  nn <- length(cliq)
  temp1 <- rep("", nrow(a))
  assignment <- c(letters, LETTERS, paste(letters, '.', sep = ''), paste(LETTERS, '.', sep = ''))
  cliq2 <- list()
  for (j in 1:nn) {
    cliq2[[j]] <- colnames(a)[cliq[[j]]]
  }
  for (ind in 1:nrow(a)) {
    tt <- which(sapply(1:length(cliq2), function(x) colnames(a)[ind] %in% cliq2[[x]]) == TRUE)
    temp1[ind] <- paste0(assignment[tt], collapse = "")
  }
  return(temp1)
}
```


# ASReml Model and Predictions
We fit the ASReml model and calculate the pairwise differences.

```{r,warning=FALSE}
# Load data
data(oats)

# Fit the ASReml model
oats.asr <- asreml(yield ~ Variety * Nitrogen, random = ~ Blocks / Wplots, data = oats)

# Perform ASReml tests
mod.asrt <- as.asrtests(oats.asr, NULL, NULL)

# Predict and calculate pairwise differences
diffs <- predictPlus(classify = "Variety*Nitrogen",
                     asreml.obj = oats.asr,
                     wald.tab = mod.asrt$wald.tab,
                     pairwise = TRUE)

# Calculate LSD at p = 0.05 level
diffs$predictions$LSD <- LSDlabels(diffs$p.differences, 0.05)
```


# Plotting Results with ggplot2
We use ```ggplot2``` to visualize the predicted values with LSD labels.

```{r,warning=FALSE}
g1 <- ggplot(data = diffs$predictions, aes(Variety, predicted.value, fill = Nitrogen)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_text(aes(Variety, predicted.value, label = round(predicted.value, 2)), 
            vjust = -0.5, size = 5, position = position_dodge(0.9)) + 
  geom_text(aes(label = LSD), vjust = 1.6, size = 5, position = position_dodge(0.9)) +
  ylim(0, 135) +
  scale_fill_brewer(palette = "Paired") +
  ggtitle("Using LSDlabels function for LSD") +
  theme_bw() +
  theme(text = element_text(size = 15), legend.position = "bottom")
g1
```


# Equivalent Method Using ```agricolae``` Package
We demonstrate an equivalent method using the ```agricolae``` package.

### Predict using predictPlus.asreml
```{r,warning=FALSE}
pred <- predictPlus.asreml(oats.asr, 
                           classify = "Variety*Nitrogen", 
                           wald.tab = as.data.frame(wald(oats.asr)), 
                           pairwise = TRUE)

# Prepare data for agricolae
std.error <- pred$predictions$standard.error
prob.matrix <- ifelse(is.na(pred$p.differences), 1, pred$p.differences)
treatments <- colnames(prob.matrix)
means <- pred$predictions$predicted.value
alpha <- 0.05

# Calculate LSD means table using agricolae
lsdmeantab <- orderPvalue(treatments, means, alpha, prob.matrix, console = TRUE)
lsdmeantab$Treatment <- rownames(lsdmeantab)

# Split Treatment column into Variety and Nitrogen
lsdmeantab <- lsdmeantab %>% 
  separate(Treatment, into = c("Variety", "Nitrogen"), sep = ",")

# Plot the results using ggplot2
g2 <- ggplot(data = lsdmeantab, aes(Variety, means, fill = Nitrogen)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_text(aes(Variety, means, label = round(means, 2)), 
            vjust = -0.5, size = 5, position = position_dodge(0.9)) + 
  geom_text(aes(label = groups), vjust = 1.6, size = 5, position = position_dodge(0.9)) +
  ylim(0, 135) +
  scale_fill_brewer(palette = "Paired") +
  ggtitle("Using agricolae for LSD") +
  theme_bw() +
  theme(text = element_text(size = 15), legend.position = "bottom")
g2
```

# Conclusion

This document demonstrates two equivalent methods for generating plots with LSD labels of ```ASReml-R``` models. The first method uses the ```LSDlabels``` function, while the second method utilises the ```agricolae``` package. Both methods produce same results, allowing users to choose the approach that best suits their needs.

```{r,warning=FALSE,fig.width=10}
grid.arrange(g1, g2, ncol = 2)
```






